\subsection{Methodology}
    We simulate and scenario corresponding to a hypothetical but plausible
initial conditions and parameters. We integrate model in
\Cref{eqn:lockdown_vaccination_ocp} by classic Runge Kutta scheme and solve the
optimization stage with the so called Differntial Evolution method.
Differential Evolution (DE) \cite{Storn1997} is an evolutionary
algorithm successfully employed for global optimization
\cite{Bilal2020}. The method is designed to optimize functions
$f:\mathbb{R}^n \to \mathbb{R}$. Nevertheless, DE can be applied to
optimize a functional as stated in \cite{CANTUNetAl}. The method can be
coded following Algorithm \ref{alg:DE1}, where an initial random
population on the search space $\mathcal{V}$ of size $N_p$ is subjected
to mutation, crossover and selection. After this process a new
population is created which, again would be subjected to the
evolutionary process. This process is repeated until some stopping
criteria is fulfilled. Finally the best individual (according to some
objective function $f_{ob}$ to optimize) is extracted. These operations
are conducted by the operators $\mathbf{X}_0$, $\mathbf{M}$,
$\mathbf{C}$, $\mathbf{S}$, $\mathbf{x}_{best}$;  whose explicit form
are coded in \cite{Penunuri2016}.

In the optimization of this study
the mutation scale factor $F$ and the crossover probability $C_r$ were
taken as 1 and 0.3 respectively, additional $N_p$ has been taken as 4
times the number of parameters (the dimension of the vector used to
describe the two controls--see \cite{CANTUNetAl}), which in our case was
of 180. As stopping criteria we have used a maximum number of
generations which is taken as \num{5000}.

    We provide in \cite{gitHub_b} a GitHub repository with all regarding R
and Fortran sources for the sake of reproductivity. This repository also
encloses data sources and a Wolfrang Mathematica notebook to reproduce all
reported figures.


\begin{algorithm}[htb]
  \caption{Differential Evolution Algorithm}
  \label{alg:DE1}
  \begin{algorithmic}
    % \small
    \State $X \leftarrow \mathbf{X}_0(Np,\mathcal{V})$
    \While{(the stopping criterion has not been met)}
    \State $M \leftarrow \mathbf{M}(X,F,\mathcal{V})$
    \State $C \leftarrow \mathbf{C}(X,M,C_r)$
    \State $X \leftarrow \mathbf{S}(X,C,f_{ob})$
    \EndWhile
    \State $\mathbf{x}_{best} \leftarrow \mathbf{Best}(X, f_{ob})$
  \end{algorithmic}
\end{algorithm}

