  %!TEX root = main.tex
\paragraph{Background}
         At the date of writing this manuscript, a Pfizer-Biotech vaccine is 
    implementing in the USA. This vaccine development among Astra-Zeneca, 
    Cansino, Sputnik, Novavax another's promises deliver sufficient doses 
    for Latinoamerica, particularly in Mexico this past Christmas has been 
    arriving the firs stock with around 40 000 amounts. In October, WHO 
    established a recommended protocol for prioritizing access to this 
    pharmaceutical hope, given clear lines about who has to be vaccinated first 
    and why.However, each vaccine development implies different issues to its 
    application. For example, the Pfizer-Biotech vaccine requires two doses and 
    very particularly logistic requirements that demand special services. In 
    Mexico, despite Pfizer taking the responsibility to capacitate and help 
    manage the immunization, we observe an explicit demand for health-logistic 
    resources that limit our institutions' response. Thus our research interest 
    in this manuscript explores the effect of the combined interventions 
    Lockdown-Vaccination to mitigate COVID-19.


\paragraph{Litterature review}
        The issue of how vaccine first has been traduced as an optimal 
    allocation problem of vaccine doses, we recommend to the interested reader 
    the articles \cite{Bubar2020,Matrajt2020}.  These articles 
    consider scenarios where the health services response and vaccine stock 
    achieve the given vaccination policy's objectives and respond to the 
    critical question of how much doses allocate to each different group 
    according to risk and age to minimize the burden of COVID-19.
    
        Early articles about COVID-19 optimal intervention modeling mainly 
    focus on Nonpharmaceutical interventions (NPIs). Mostly these works 
    understand the control strategy as the diminish of contact rates by 
    reducing mobility or modulating parameters regarding the generation of new 
    infections by linear controls \cite{Naraigh2020,Ullah2020}, 
    Lockdown-Quarantine \cite{Mandal2020},  shield immunity 
    \cite{Weitz2020}. Libotte et. al. reports in \cite{Libotte2020} optimal 
    vaccination strategies for COVID19.
\paragraph{Contribution and main objectives}
        Our manuscript is the first contribution modeling with optimal control 
    of Lockdown-Vaccination strategies' effect to the best of our knowledge. 
    Since health services' response will be limited by the vaccine stock
    and logistics, to implement in parallel NPIs is mandatory. We focus on 
    formulating and studying via simulation the system Lockdown-Vaccination 
    with recent and approved vaccine profile by the  Mexico Health council and 
    developing optimal policies for the Lockdown release-input and Vaccine 
    application doses.
\paragraph{Vaccine development}
        According to official Governmental communication in December, Mexico 
    treated  \SI{36000000}{doses} Pfizer-Biotech, \SI{76000000}{doses} with 
    Aztra-Seneca \SI{18000000}{doses} of Cansino-BIO. Other developments 
    also are running the  third Phase, and with high probability,  in the 
    third quarter of 2021, some of these developments will incorporate into 
    Mexico's vaccine portfolio. Despite official agreements, each vaccine's 
    delivery schedule is under uncertainty and-or subject to the approval 
    of COFEPRIS.
\paragraph{Problem setup}
        The first accepted vaccine \textemdash Pfizer-BioNTech's BNT162b2 
    \textemdash has an efficacy above \SI{90}{\percent}  and requires 
    two doses to achieve immunity. The other mentioned developments have a very 
    similar profile but require different logistic management and stock 
    allocation.  Thus, we face designing a schedule of dose application subject 
    to a given vaccine stock that will be applied in a given period. To this 
    end, we formulate an optimal control problem that minimizes the burden of 
    COVID-19 in DALYs [WhoDALY(2020)]. We also optimize the cost generated by 
    the implementation of Vaccination in parallel with Lockdown.

\paragraph{Piecewise optimal policies}
        Comment about the solution of the underlying Optimal Control Problem
    \comment[id=SDIV]{David}
    One of the main features of our model is that we consider piecewise 
    constant control policies instead of general measurable control policies 
    (also called permanent controls) to minimize a cost functional. General 
    control policies are difficult to implement since the authority has to make 
    different choices every permanently. The optimal policies we find are 
    constant in each interval of time and hence these policies are easier to 
    implement. 
        Optimal control problems with piecewise constant policies have been 
    studied in different contexts. For instance, a solution method based on the 
    gradient of the cost functional is studied in \cite{MR3223602}; convergence 
    results of piecewise constant solutions to permanent solutions in 
    linear-quadratic problems are given in \cite{MR3627992}; or, in 
    \cite{CANTUNetAl}, a general numerical methodology to find piecewise 
    constant solutions is proposed. 
\paragraph{Papaer structure}